Data Specialization for a Staged Lambda Calculus

### 2015-02-20 paper/implementation.tex

\begin{abstrsyn}
We have implemented the semantics and splitting algorithm of \lang\ in Standard ML,
as well as a semantics for the output language.
In addition to the features shown here, our implementation includes base types, 
$n$-ary products and sums, patterns, if and let statements, and a basic elaboration step.
We cover some of the details here.

\paragraph {Splitting Optimizations} 
Applying the rules of \cref{sec:splitting} naively can yield inefficient split programs.
In many cases, the results can be improved by considering special cases in individual rules.
For instance, trivial subterm precomputations needn't be saved at tuples:
\[
\infer {\splittwo {\tup{e_1,e_2}} A p l {\tup{r_1,r_2}}} {\splittwo {e_1} A {\tup{}} {\_}{r_1} & \splittwo {e_2} A  p l {r_2}}
\]
Our implementation uses many optimizations like these.

\paragraph {Patterns}

Our implemention includes patterns to eliminate products, $\fut$, and $\curr$.
We split patterns directly (in a way similar to the standard elim forms), 
rather than translate them away before splitting.
This, along with the optimized splitting algorithm,
goes a long way toward producing readable outputs.

\paragraph {Syntax}

Our concrete syntax is used in examples throughout the paper.
Its main notable property is that world boundaries are always noted by braces.

@@

>>>>

\begin{abstrsyn}
We have implemented the semantics and splitting algorithm of \lang\ in Standard ML.
Specifically, our implemention includes parsers, printers, and interpretters for both \lang\ 
and the output language, as well as a splitting algorithm which mediates between the two.
In addition to the features shown here, our implementation includes base types, 
$n$-ary products and sums, patterns, if and let statements, and a basic elaboration step.
We cover some of the details here.

\paragraph {Syntax}
Our concrete syntax is used in examples throughout the paper.
Its main notable property is that world boundaries are always noted by braces.

@@

\paragraph {Patterns}
Our implemention includes patterns in the input language to eliminate products, $\fut$, and $\curr$.
As implied by our splitting rules, we also include product patterns in the output language.
For the sake of readability, we split input patterns directly (in a way similar to the standard elim forms), 
rather than translate them away before splitting.

\paragraph {Specialized Splitting} 
Our implemented splitting algorithm also includes many special cases to create more readable outputs.
Most important among these are cases for eliminating trivial precomputations.
These special cases all fit within the presented framework of our correctness condition.
